---
title: "Take a Hike on the Perfect Trail for You!"
author: Natalia Iannucci, Hana Hirano, and Syrine Matoussi
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE}
library(tidyverse)
library(macleish)
library(leaflet)
library(sf)
library(ggthemes)
library(ggrepel)

boundary <- macleish_layers %>% 
  pluck("boundary")
macleish_trails <- macleish_layers %>%
  pluck("trails")

interior_trails <- st_intersection(
  boundary, 
  macleish_trails
)

trails_leaflet <- leaflet() %>%
  addTiles() %>%
  addPolylines(data = interior_trails) 

trails_leaflet
```

```{r message=FALSE}
elevation <- macleish_layers %>%
  pluck("contours_3m") %>%
  st_transform(4326)
```

```{r message=FALSE}
elevation_trails <- st_intersection(
  interior_trails,
  elevation
) 

```

```{r message=FALSE}
grouped <- elevation_trails %>%
  filter(name %in% c("entry trail", "Western Loop", "Eastern Loop", "Vernal Pool Loop", "Porcupine Trail", "Easy Out")) %>%
  group_by(name) %>%
  summarize(elevation_ft = sum(ELEV_FT), distance = sum(SHAPE_LEN))

grouped <- grouped %>% 
  mutate("Distance_miles" = distance / 5280) %>%
  mutate("Difficulty" = sqrt(elevation_ft * 2 * Distance_miles))
ggplot(data= grouped, aes(y = elevation_ft, x = Distance_miles)) +geom_line() + geom_area(color="black", fill="gray" )+labs(title= "elevation profiles at MacLeish",face="bold", size=25 )+labs(y= "feet above sea level")+labs(x="Miles")+ geom_text(aes(label = name), size = 3)
```

```{r message=FALSE}
pal <- colorNumeric(c("Black", "Red", "Blue", "Yellow", "Pink", "Purple"), 1:6)

#Ideal -> (Yellow = entry trail)
#Red = Western Loop (+ entry trail, and the tiny one on the top, and part of the straight-ish poplar hill road)(right now)
#Blue = Porcupine Trail (+ the bottom part of Western Loop, and snowmobile trail on the bottom right)
#Black = Vernal Pool Loop
#Pink = Eastern Loop = Homestead Trail? = 5
#Purple = Easy Out (where is it, I don't know which one is Easy Out..., so it's not illustrated)

difficulty_map <- leaflet() %>%
  addTiles() %>%
  addPolylines(data = interior_trails, color = ~pal(c(3, 2, 5, 1, 5, 2)))%>%
  addMarkers(data = st_cast(trail_stream_xings, "POINT"), popup = ~name)

difficulty_map
```

```{r message=FALSE}
a  <- elevation_trails %>%
  filter(name == "Western Loop") %>%
  filter(ELEV_FT == "797")

leaflet() %>%
  addTiles() %>%
  addPolylines(data = a)
```

```{r message=FALSE}
elevation_graph <- ggplot(grouped, aes(x = Distance_miles, y = elevation_ft)) +
  geom_point(size = 5) +
  theme_solarized() +
  geom_smooth(color = "black", fill = "blue") +
  geom_label_repel(aes(label = name), point.padding = 0.5, hjust = "middle", vjust = "center") +
  xlab("Distance of Trail (in miles)") +
  ylab("Total change in elevation (in feet)") 
```

```{r message=FALSE}  
elevation_graph
```


SQRT(Elevation Gain(in feet) x 2 x distance (in miles))
- less than 50 : Easiest (1.5 miles per hour)
- 50 ~ 100 : Moderate (1.4 miles per hour)
- 100 ~ 150 : Moderately Strenuous (1.3 mph)
- 150 ~ 200 : Strenuous (1.2 mph)
- 200 ~ : Very Strenuous (1.2 mph)


Our goal is to determine the difficulty of each trail at MacLeish based on its length and elevation, and to be able to classify trails based on difficulty so that visitors can better decide which trail they would like to take. ^[https://github.com/niannucci/MP3-Macleish]

> Word count: `r wordcountaddin::word_count()`